Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality

### 2015-06-18 content/contribution/continuous-datalog.tex
slight modifications to make the nature of entailment instances consistent

	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
	\begin{description}
		\item[Instant\label{def:continuous datalog: instant}]\hfill\\
			An instant is a discrete point in time.
			The continuous distance in time between these discrete points may be theoretically infinitesimal, but will in practise have a lower bound set by hardware limitations, or will be defined as part of some specific implementation.
			In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
		\item[Stream of Facts\label{def:continuous datalog: stream}]\hfill\\
			A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g. the time at which each was entailment by some process at the source of the stream.
			Each instance of a fact is considered to be uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
			This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
			In practice, if two instances of the same fact also share an entailment time, they are considered to be the same instance.
		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
			A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
			The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
	\end{description}
	The following axioms define the semantics of streams and windows over those streams in Continuous Datalog.
	\begin{enumerate}\newcounter{continuousDatalogAxioms}
		\item\label{axiom:continuous datalog: window range leq 0}
			A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
			\begin{equation*}
				W^{S,X}_{i} = \varnothing
			\end{equation*}
		\item\label{axiom:continuous datalog: window composition}
			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
			\begin{equation*}
				W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
			\end{equation*}
		\item\label{axiom:continuous datalog: window disjointness}
			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
			\begin{equation*}
				\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
			\end{equation*}
		% \item\label{axiom:continuous datalog: stream disjointness}
		% 	The contents of any two streams are disjoint, thus a window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over any other stream ${Z \neq S}$ at any instant ${i \in \mathbb{N}}$.
		% 	\begin{equation*}
		% 		W^{S,X}_{i} \cap W^{Z,Y}_{i} = \varnothing
		% 	\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

@@

	From this, we can see that each instance $e$ of an entailment in some $E_{i}$ may take the same form as any other fact in a stream of facts, being ${e = p(s,o):t}$, where $t$ is the real time associated with the instant in which $e$ is first entailed by the ${CDP}$ (i.e. it appears in $E_{i}$ but not $E_{i-1}$, and so is part of $W^{P,1}_{i}$).
	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time of the entailments but by the time at which they \emph{cease} to be entailed (i.e. they appear in $E_{i-1}$ but not $E_{i}$, and so are part of $W^{N,1}_{i}$).

	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
	However, it is proven in \refapp{semantics: proof for entailments} that, from these axioms, it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
	The minimum range $X$ is such to achieve the \emph{complete} set of entailment instances $E_{i}$ of the ${CDP}_{i}$.
	In addition, the minimum range $Y$ is such to achieve the \emph{correct} set $E_{i}$.
	As such, a \emph{valid window} that contains only and exactly those instances of entailments in $E_{i}$ at every instant $i$ may be trivially maintained from the streams $P$ and $N$ continuously entailed by some ${CDP}$.
\end{nestedsection}

>>>>

	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
	\begin{description}
		\item[Instant\label{def:continuous datalog: instant}]\hfill\\
			An instant is a discrete point in time.
			The continuous distance in time between these discrete points may be theoretically infinitesimal, but will in practise have a lower bound set by hardware limitations, or will be defined as part of some specific implementation.
			In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
		\item[Stream of Facts\label{def:continuous datalog: stream}]\hfill\\
			A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g. the time at which each was entailment by some process at the source of the stream.
			Each instance of a fact is uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
			This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
			A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
			The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
	\end{description}
	The following axioms define the semantics of streams and windows over those streams in Continuous Datalog.
	\begin{enumerate}\newcounter{continuousDatalogAxioms}
		\item\label{axiom:continuous datalog: window range leq 0}
			A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
			\begin{equation*}
				W^{S,X}_{i} = \varnothing
			\end{equation*}
		\item\label{axiom:continuous datalog: window composition}
			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
			\begin{equation*}
				W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
			\end{equation*}
		\item\label{axiom:continuous datalog: window disjointness}
			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
			\begin{equation*}
				\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
			\end{equation*}
		% \item\label{axiom:continuous datalog: stream disjointness}
		% 	The contents of any two streams are disjoint, thus a window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over any other stream ${Z \neq S}$ at any instant ${i \in \mathbb{N}}$.
		% 	\begin{equation*}
		% 		W^{S,X}_{i} \cap W^{Z,Y}_{i} = \varnothing
		% 	\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

@@

	From this, we can see that each instance of an entailment in some $E_{i}$ may take the same form as an instance of any other fact in a stream of facts, being ${p(s,o):t}$, where $t$ is the real time associated with the instant $i_{e}$ in which ${p(s,o)}$ becomes entailed by the ${CDP}$
	(i.e. when it becomes newly justified by some subset of instances in $W^{S,R}_{i_{e}}$).
	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time $i_{e}$ of the entailment instances but by the time at which their justification \emph{ceases} to be valid
	(i.e. the first instant ${i_{n} \geq i_{e}}$ in which $W^{S,R}_{i_{n}}$ no longer contains the subset of fact instances first present in $W^{S,R}_{i_{e}}$ that justified the entailment instance ${p(s,o):t}$).

	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
	However, it is proven in \ref{sec:semantics: proof for entailments}, from these axioms, that it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
	The minimum range $X$ is such to achieve the \emph{complete} set of entailment instances $E_{i}$ of the ${CDP}_{i}$.
	In addition, the minimum range $Y$ is such to achieve the \emph{correct} set $E_{i}$.
	As such, a \emph{valid window} that contains only and exactly those instances of entailments in $E_{i}$ at every instant $i$ may be trivially maintained from the streams $P$ and $N$ continuously entailed by some ${CDP}$.
\end{nestedsection}
